---
title: ""
pagetitle: "M252 Dope"
output: 
  html_document:
    theme: yeti
    toc: false
    toc_float: true
    toc_depth: 4
    css: ['../../styles/rpruim2020.css', '../../styles/folding.css']
    includes:
      before_body: ../header.html
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(qrcode)
library(mosaic)
library(rjp)
library(tidyr)
set.seed(123)
spaces <- function(n) {
  paste(rep(" ", n), collapse = "")
}
Roster <- read.csv('../../../grades/S21/Math252-S21-roster-clean.csv') %>%
  mutate(name = paste(first, last))
```

<script src="../../styles/folding.js"></script>

<style>
div.indent {
  margin: 0 0 0 25px;
}
</style>

\def\prob{\operatorname{P}}
\def\Prob{\operatorname{P}}
\def\tand{\operatorname{and}}
\def\tor{\operatorname{or}}
\def\tif{\operatorname{if}}
\def\intersect{\operatorname{and}}
\def\union{\operatorname{or}}
\def\intersect{\cap}
\def\union{\cup}
\def\To{\Rightarrow}
\def\emptystring{\lambda}


# Dope Sheets {.tabset}

**dope**[^webster]

* *n.* information, especially from a reliable source [the inside dope];  
* *v.* figure out -- usually used with out;                              
* *adj.* excellent

[^webster]: Definitions selected from Webster's online dictionary  

Here's where you can come to find out what's up each day. Homework 
assignments are available on the [**homework page**](../hw/). 

And here is a link to the 
[**Progress Log**](https://docs.google.com/spreadsheets/d/1uPQf12t707jXYaEEGgfD5jWbOhp5EjynERKDuNIji0E/edit?usp=sharing) 
we will often use when working in groups on Teams.

[Dope sheets for weeks 1 - 5](dope01.html)

[Dope sheets for weeks 6 - 7](dope02.html)

[Dope sheets for weeks 8 - 9](dope03.html)


## Week 10 {.active}

### Mon 4/5

* New Groups

```{r echo = FALSE, results = "asis", include = TRUE}
set.seed(40520215)
R <- shuffle(Roster) %>% group_by(section) %>% 
  mutate(group = rep(1:5, length.out = n())) %>%
  mutate(group = ifelse(section == "B", group + 5, group))
         
for(g in 1:10) {
  cat(paste0('  * Group ', g, ": ", paste(R %>% filter(group == g) %>% pull(name) %>% sort(), collapse = ", "), "\n"))
}
```

* Homework

  * Find homework in the "after test 2" section now.
  * PS 13 due on Wednesday

* **Topic of the Day:** More finite state automata 

  * Worksheet: 
    [[HTML]](models-of-computation/finite-state-automata.html)
    [[PDF]](models-of-computation/finite-state-automata.pdf)
    
  * **Start at question 3** to get back into this after a week away from
  automata.

### Wed 4/7

* **Topic of the Day:** DFAs and NFAs

  * A couple things from last time
  
      * #6 d and e -- how to design a DFA (and what the states are) 
      * definition of NFA
      
  * Worksheet: 
    [[HTML]](models-of-computation/finite-state-automata.html)
    [[PDF]](models-of-computation/finite-state-automata.pdf)
    
      * Start at #10 and don't go beyond #19.
      * If you get through #19, don't leave without checking in with me.

### Fri 4/9


