---
title: ""
pagetitle: "M252 Dope"
output: 
  html_document:
    theme: yeti
    toc: false
    toc_float: true
    toc_depth: 4
    css: ['../../styles/rpruim2020.css', '../../styles/folding.css']
    includes:
      before_body: ../header.html
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(qrcode)
library(mosaic)
library(rjp)
library(tidyr)
set.seed(123)
spaces <- function(n) {
  paste(rep(" ", n), collapse = "")
}
Roster <- read.csv('../../../grades/S21/Math252-S21-roster-clean.csv') %>%
  mutate(name = paste(first, last))
```

<script src="../../styles/folding.js"></script>

<style>
div.indent {
  margin: 0 0 0 25px;
}
</style>

\def\prob{\operatorname{P}}
\def\Prob{\operatorname{P}}
\def\tand{\operatorname{and}}
\def\tor{\operatorname{or}}
\def\tif{\operatorname{if}}
\def\intersect{\operatorname{and}}
\def\union{\operatorname{or}}


# Dope Sheets {.tabset}

**dope**[^webster]

* *n.* information, especially from a reliable source [the inside dope];  
* *v.* figure out -- usually used with out;                              
* *adj.* excellent

[^webster]: Definitions selected from Webster's online dictionary  

Here's where you can come to find out what's up each day. Homework 
assignments are available on the [**homework page**](../hw/). 

And here is a link to the 
[**Progress Log**](https://docs.google.com/spreadsheets/d/1uPQf12t707jXYaEEGgfD5jWbOhp5EjynERKDuNIji0E/edit?usp=sharing) 
we will often use when working in groups on Teams.

[Dope sheets for weeks 1 - 5](dope01.html)

[Dope sheets for weeks 6 - 7](dope02.html)


## Week 8 {.active}

### Mon 3/22

* **Test 2** will be on Friday, April 2. I had no requests for that date to be a black out date.
More details to come.

* **Topic of the Day:** More Expected Value


  * Worksheet: 
  [[HTML]](counting-and-probability/more-expected-value.html)
  [[PDF]](counting-and-probability/more-expected-value.pdf)

<div class = "indent">
<div class = "explain" label = "Hints and Solutions">

<div class = "explain" label = "Exercise 10.2">
value        |    0  |   1  
:----------: | :---: | :---: 
probability  |  0.5  |  0.5
</div>

<div class = "explain" label = "Exercise 10.3">
The table for $X^2$ is exactly the same as the table for $X$ in this case. (That's unusual,
do you see when this will happen?)
</div>

<div class = "explain" label = "Exercise 10.4">

* $\Prob(X = 3 \land  Y = 2) = 0.10$. 
* $\Prob(X = 3) = 0.25 + 0.10 + 0.05 = 0.4$.  
* $\Prob(Y = 2) = 0.10 + 0.05 + 0.10 = 0.25$.
</div>

<div class = "explain" label = "Exercise 10.6">


value of $X$ |   1    |   2    |    3
-------------| :----: | :----: | :----:
probability  |  0.30  |  0.30  |  0.40

value of $Y$ |   1    |   2    |    3
-------------| :----: | :----: | :----:
probability  |  0.45  |  0.25  |  0.30

value of $X + Y$ |   2    |   3    |    4   |   5    |   6   
-----------------| :----: | :----: | :----: | :----: | :----: 
probability      |  0.10  |  0.20  |  0.40  |  0.25  |  0.05

</div>

<div class = "explain" label = "Exercise 10.8">

a) 

* $E(3 X + 2) = 3 E(X) + 2 = 3 \cdot 20 + 2 = 62$.
* $E(2 X − 3) = 2 E(X) - 3 = 2\cdot 20 - 3 = 37$.

b) 

* $E (X +Y) = 5 + 2 = 7$; $E(X - Y) = E(X + (-Y)) = 5 + (-2) = 3$

</div>

<div class = "explain" label = "Exercise 10.10">
$E(X_1 + X_2) = E(X_1) + E(X_2) = 3.5 + 3.5 = 7$.
</div>

<div class = "explain" label = "Exercise 10.11">
The expected value is the same no matter how many people there are.
If you didn't get that, double check your work.
</div>

<div class = "explain" label = "Exercise 10.12">
Answer: $p$
</div>

<div class = "explain" label = "Exercise 10.13">
Answer: $np$
</div>

<div class = "explain" label = "Exercise 10.15">
Hint: What is the expected number of diamonds on the first card?
</div>

<div class = "explain" label = "Exercise 10.16">
Hint: Can you split this into a sum of 5 terms?
</div>

<div class = "explain" label = "Exercise 10.17">
Hint: Let $X_i = 1$ if the number $i$ is rolled and 0 other wise.  
What is $\E(X_i)$ and how is it related to the main question?
</div>
</div>
</div>

### Wed 3/24

* New Groups

```{r echo = FALSE, results = "asis", include = TRUE}
set.seed(3242021)
R <- shuffle(Roster) %>% group_by(section) %>% 
  mutate(group = rep(1:5, length.out = n())) %>%
  mutate(group = ifelse(section == "B", group + 5, group))
         
for(g in 1:10) {
  cat(paste0('  * Group ', g, ": ", paste(R %>% filter(group == g) %>% pull(name) %>% sort(), collapse = ", "), "\n"))
}
```

### Fri 3/26
