---
title: ""
pagetitle: "M252 Dope"
output: 
  html_document:
    theme: yeti
    toc: false
    toc_float: true
    toc_depth: 4
    css: ['../../styles/rpruim2020.css', '../../styles/folding.css']
    includes:
      before_body: ../header.html
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(qrcode)
library(mosaic)
library(rjp)
library(tidyr)
set.seed(123)
spaces <- function(n) {
  paste(rep(" ", n), collapse = "")
}
Roster <- read.csv('../../../grades/S21/Math252-S21-roster-clean.csv') %>%
  mutate(name = paste(first, last))
```

<script src="../../styles/folding.js"></script>


\def\prob{\operatorname{P}}
\def\Prob{\operatorname{P}}
\def\tand{\operatorname{and}}
\def\tor{\operatorname{or}}
\def\tif{\operatorname{if}}
\def\intersect{\operatorname{and}}
\def\union{\operatorname{or}}


# Dope Sheets {.tabset}

**dope**[^webster]

* *n.* information, especially from a reliable source [the inside dope];  
* *v.* figure out -- usually used with out;                              
* *adj.* excellent

[^webster]: Definitions selected from Webster's online dictionary  

Here's where you can come to find out what's up each day. Homework 
assignments are available on the [**homework page**](../hw/). 

And here is a link to the 
[**Progress Log**](https://docs.google.com/spreadsheets/d/1uPQf12t707jXYaEEGgfD5jWbOhp5EjynERKDuNIji0E/edit?usp=sharing) 
we will often use when working in groups on Teams.

## Week 1 

### Wed 2/3

* Welcome to Math 252!

* **Topic of the Day**: Intro to Graphs

  * Worksheet:  [[PDF]](graphs/intro-to-graphs.pdf)
  
  <!-- [[HTML]](graphs/introduction-to-graphs.html)  -->
  
* [Homework](../hw/) due Friday

  * Officially due at 11:59 pm, but you should attempt to have it done before class.
  * Submit via [Gradescope](https:://gradescope.com) -- details on [the homework page](../hw)

  
### Fri 2/5

* Meet in Teams

* Groups

```{r echo = FALSE, results = "asis", include = TRUE}
set.seed(205)
R <- shuffle(Roster) %>% group_by(section) %>% 
  mutate(group = rep(1:5, length.out = n())) %>%
  mutate(group = ifelse(section == "B", group + 5, group))
         
for(g in 1:10) {
  cat(paste0('  * Group ', g, ": ", paste(R %>% filter(group == g) %>% pull(name), collapse = ", "), "\n"))
}
```

* **Topic of the Day:** More about Graphs

  * Worksheet: [[PDF]](graphs/more-graphs.pdf)
  
    * If you print this out, I recommend printing 1-sided, or printing the first page separate from the other two.
    That will make it easier to work with (less page flipping).
    
    * Keep track of your group's progress with this 
    [Progress Log](https://docs.google.com/spreadsheets/d/1uPQf12t707jXYaEEGgfD5jWbOhp5EjynERKDuNIji0E/edit?usp=sharing) 
     

## Week 2 {.active}

### Mon 2/8

* Meet in Teams

* Same groups as Friday

* **Topic of the Day**: Representing Graphs

  * Worksheet: 
  [[PDF](graphs/graph-representations.pdf)]
  [[HTML](graphs/graph-representations.html)]
  
* WebWork assignments due tonight and Wednesday night

### Wed 2/10

* Meet in Teams

* Same groups as Friday

* **Topic of the Day**: Graph Isomorphism

  * Worksheet: 
  [[PDF](graphs/graph-isomorphism.pdf)]
  [[HTML](graphs/graph-isomorphism.html)]
  
* WW2 due tonight, PS 2 on Monday.


### Fri 2/12
