---
title: ""
pagetitle: "M252 Dope"
output: 
  html_document:
    theme: yeti
    toc: false
    toc_float: true
    toc_depth: 4
    css: ['../../styles/rpruim2020.css', '../../styles/folding.css']
    includes:
      before_body: ../header.html
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(qrcode)
library(mosaic)
library(rjp)
library(tidyr)
set.seed(123)
spaces <- function(n) {
  paste(rep(" ", n), collapse = "")
}
Roster <- read.csv('../../../grades/S21/Math252-S21-roster-clean.csv') %>%
  mutate(name = paste(first, last))
```

<script src="../../styles/folding.js"></script>

<style>
div.indent {
  margin: 0 0 0 25px;
}
</style>

\def\prob{\operatorname{P}}
\def\Prob{\operatorname{P}}
\def\tand{\operatorname{and}}
\def\tor{\operatorname{or}}
\def\tif{\operatorname{if}}
\def\intersect{\operatorname{and}}
\def\union{\operatorname{or}}


# Dope Sheets {.tabset}

**dope**[^webster]

* *n.* information, especially from a reliable source [the inside dope];  
* *v.* figure out -- usually used with out;                              
* *adj.* excellent

[^webster]: Definitions selected from Webster's online dictionary  

Here's where you can come to find out what's up each day. Homework 
assignments are available on the [**homework page**](../hw/). 

And here is a link to the 
[**Progress Log**](https://docs.google.com/spreadsheets/d/1uPQf12t707jXYaEEGgfD5jWbOhp5EjynERKDuNIji0E/edit?usp=sharing) 
we will often use when working in groups on Teams.

## Week 1 

### Wed 2/3

* Welcome to Math 252!

* **Topic of the Day**: Intro to Graphs

  * Worksheet:  [[PDF]](graphs/intro-to-graphs.pdf)
  
  <!-- [[HTML]](graphs/introduction-to-graphs.html)  -->
  
* [Homework](../hw/) due Friday

  * Officially due at 11:59 pm, but you should attempt to have it done before class.
  * Submit via [Gradescope](https:://gradescope.com) -- details on [the homework page](../hw)

  
### Fri 2/5

* Meet in Teams

* Groups

```{r echo = FALSE, results = "asis", include = TRUE}
set.seed(205)
R <- shuffle(Roster) %>% group_by(section) %>% 
  mutate(group = rep(1:5, length.out = n())) %>%
  mutate(group = ifelse(section == "B", group + 5, group))
         
for(g in 1:10) {
  cat(paste0('  * Group ', g, ": ", paste(R %>% filter(group == g) %>% pull(name), collapse = ", "), "\n"))
}
```

* **Topic of the Day:** More about Graphs

  * Worksheet: [[PDF]](graphs/more-graphs.pdf)
  
    * If you print this out, I recommend printing 1-sided, or printing the first page separate from the other two.
    That will make it easier to work with (less page flipping).
    
    * Keep track of your group's progress with this 
    [Progress Log](https://docs.google.com/spreadsheets/d/1uPQf12t707jXYaEEGgfD5jWbOhp5EjynERKDuNIji0E/edit?usp=sharing) 
     

## Week 2 {.active}

### Mon 2/8

* Meet in Teams

* Same groups as Friday

* **Topic of the Day**: Representing Graphs

  * Worksheet: 
  [[PDF](graphs/graph-representations.pdf)]
  [[HTML](graphs/graph-representations.html)]
  
* WebWork assignments due tonight and Wednesday night

### Wed 2/10

* Meet in Teams

* Same groups as Friday

* **Topic of the Day**: Graph Isomorphism

  * Worksheet: 
  [[PDF](graphs/graph-isomorphism.pdf)]
  [[HTML](graphs/graph-isomorphism.html)]
  
* WW2 due tonight, PS 2 on Monday.

<div class = "indent">
<div class = "explain" label = "Show some comments on the worksheet">
In problem 2, the two matrices are **not** graph invariants.  The problem is
that for the matrices to match, the order of the vertices (and edges in the case
of an incidence matrix) must line up according to the isomorphism. If they don't,
you may have different matrices even though the graphs are isomorphic.  (In fact, you
can get different matrices for the same graph if you just change the order of the vertices.)

In problem 3, here is one way to show the second pair are not isomorphic:
Consider $x$ and $y$.  They are adjacent and each has degree 2.  But there is no pair 
of adjacent vertices with degree 2 in graph G. (There are other ways to show that these 
graphs are not isomorphic as well.)

In problem 4, 

* The second pair are directed graphs.  Be sure to make the directions match when looking
for an isomorphism.  In-degree and out-degree can be useful for directed graphs.

* The third pair is challenging.  Both graphs are 3-regular (meaning every vertex has
degree 3, so degree alone doesn't help you).
</div>
</div>
<br>

### Fri 2/12

* New Groups

```{r echo = FALSE, results = "asis", include = TRUE}
set.seed(212)
R <- shuffle(Roster) %>% group_by(section) %>% 
  mutate(group = rep(1:5, length.out = n())) %>%
  mutate(group = ifelse(section == "B", group + 5, group))
         
for(g in 1:10) {
  cat(paste0('  * Group ', g, ": ", paste(R %>% filter(group == g) %>% pull(name), collapse = ", "), "\n"))
}
```

* Reminder: 
You can request that we not have a test on a particular day by filling out 
[**this form**](https://docs.google.com/forms/d/e/1FAIpQLSdZxGUTSv8ie4itR0ErnRDpfEVtxWS5t2noU0cxnye0GgvVbQ/viewform?usp=sf_link).

* **Topic of the Day**: Paths in Graphs

  * Worksheet: 
  [[PDF](graphs/paths-in-graphs.pdf)]
  [[HTML](graphs/ipaths-in-graphs.html)]

* PS 2 due tonight; PS 3 due Monday -- Submit via GradeScope

```{r, include = FALSE}
library(networkD3)
G <- data.frame(
  from = rep(1:10, each = 3),
  to = c(2,5,10, 1,3,9, 2,8,4, 3,7,5, 1,4,6, 5,8,9, 4,6,9, 6,3,10, 
         2,6,7, 1,8,6)
)
simpleNetwork(G, height="100px", width="100px", charge = 0)
```
