---
title: ""
pagetitle: "M252 Dope"
output: 
  html_document:
    theme: yeti
    toc: false
    toc_float: true
    toc_depth: 4
    css: ['../../styles/rpruim2020.css', '../../styles/folding.css']
    includes:
      before_body: ../header.html
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(qrcode)
library(mosaic)
library(rjp)
library(tidyr)
set.seed(123)
spaces <- function(n) {
  paste(rep(" ", n), collapse = "")
}
Roster <- read.csv('../../../grades/S21/Math252-S21-roster-clean.csv') %>%
  mutate(name = paste(first, last))
```

<script src="../../styles/folding.js"></script>

<style>
div.indent {
  margin: 0 0 0 25px;
}
</style>

\def\prob{\operatorname{P}}
\def\Prob{\operatorname{P}}
\def\tand{\operatorname{and}}
\def\tor{\operatorname{or}}
\def\tif{\operatorname{if}}
\def\intersect{\operatorname{and}}
\def\union{\operatorname{or}}


# Dope Sheets {.tabset}

**dope**[^webster]

* *n.* information, especially from a reliable source [the inside dope];  
* *v.* figure out -- usually used with out;                              
* *adj.* excellent

[^webster]: Definitions selected from Webster's online dictionary  

Here's where you can come to find out what's up each day. Homework 
assignments are available on the [**homework page**](../hw/). 

And here is a link to the 
[**Progress Log**](https://docs.google.com/spreadsheets/d/1uPQf12t707jXYaEEGgfD5jWbOhp5EjynERKDuNIji0E/edit?usp=sharing) 
we will often use when working in groups on Teams.

## Week 1 

### Wed 2/3

* Welcome to Math 252!

* **Topic of the Day**: Intro to Graphs

  * Worksheet:  [[PDF]](graphs/intro-to-graphs.pdf)
  
  <!-- [[HTML]](graphs/introduction-to-graphs.html)  -->
  
* [Homework](../hw/) due Friday

  * Officially due at 11:59 pm, but you should attempt to have it done before class.
  * Submit via [Gradescope](https:://gradescope.com) -- details on [the homework page](../hw)

  
### Fri 2/5

* Meet in Teams

* Groups

```{r echo = FALSE, results = "asis", include = TRUE}
set.seed(205)
R <- shuffle(Roster) %>% group_by(section) %>% 
  mutate(group = rep(1:5, length.out = n())) %>%
  mutate(group = ifelse(section == "B", group + 5, group))
         
for(g in 1:10) {
  cat(paste0('  * Group ', g, ": ", paste(R %>% filter(group == g) %>% pull(name), collapse = ", "), "\n"))
}
```

* **Topic of the Day:** More about Graphs

  * Worksheet: [[PDF]](graphs/more-graphs.pdf)
  
    * If you print this out, I recommend printing 1-sided, or printing the first page separate from the other two.
    That will make it easier to work with (less page flipping).
    
    * Keep track of your group's progress with this 
    [Progress Log](https://docs.google.com/spreadsheets/d/1uPQf12t707jXYaEEGgfD5jWbOhp5EjynERKDuNIji0E/edit?usp=sharing) 
     

## Week 2 

### Mon 2/8

* Meet in Teams

* Same groups as Friday

* **Topic of the Day**: Representing Graphs

  * Worksheet: 
  [[PDF](graphs/graph-representations.pdf)]
  [[HTML](graphs/graph-representations.html)]
  
* WebWork assignments due tonight and Wednesday night

### Wed 2/10

* Meet in Teams

* Same groups as Friday

* **Topic of the Day**: Graph Isomorphism

  * Worksheet: 
  [[PDF](graphs/graph-isomorphism.pdf)]
  [[HTML](graphs/graph-isomorphism.html)]
  
* WW2 due tonight, PS 2 on Monday.

<div class = "indent">
<div class = "explain" label = "Show some comments on the worksheet">
In problem 2, the two matrices are **not** graph invariants.  The problem is
that for the matrices to match, the order of the vertices (and edges in the case
of an incidence matrix) must line up according to the isomorphism. If they don't,
you may have different matrices even though the graphs are isomorphic.  (In fact, you
can get different matrices for the same graph if you just change the order of the vertices.)

In problem 3, here is one way to show the second pair are not isomorphic:
Consider $x$ and $y$.  They are adjacent and each has degree 2.  But there is no pair 
of adjacent vertices with degree 2 in graph G. (There are other ways to show that these 
graphs are not isomorphic as well.)

In problem 4, 

* The second pair are directed graphs.  Be sure to make the directions match when looking
for an isomorphism.  In-degree and out-degree can be useful for directed graphs.

* The third pair is challenging.  Both graphs are 3-regular (meaning every vertex has
degree 3, so degree alone doesn't help you).
</div>
</div>
<br>

### Fri 2/12

* New Groups

```{r echo = FALSE, results = "asis", include = TRUE}
set.seed(212)
R <- shuffle(Roster) %>% group_by(section) %>% 
  mutate(group = rep(1:5, length.out = n())) %>%
  mutate(group = ifelse(section == "B", group + 5, group))
         
for(g in 1:10) {
  cat(paste0('  * Group ', g, ": ", paste(R %>% filter(group == g) %>% pull(name), collapse = ", "), "\n"))
}
```

* Reminder: 
You can request that we not have a test on a particular day by filling out 
[**this form**](https://docs.google.com/forms/d/e/1FAIpQLSdZxGUTSv8ie4itR0ErnRDpfEVtxWS5t2noU0cxnye0GgvVbQ/viewform?usp=sf_link).

* **Topic of the Day**: Paths in Graphs

  * Worksheet: 
  [[PDF](graphs/paths-in-graphs.pdf)]
  [[HTML](graphs/paths-in-graphs.html)]

* PS 2 due tonight; PS 3 due Monday -- Submit via GradeScope

```{r, include = FALSE}
library(networkD3)
G <- data.frame(
  from = rep(1:10, each = 3),
  to = c(2,5,10, 1,3,9, 2,8,4, 3,7,5, 1,4,6, 5,8,9, 4,6,9, 6,3,10, 
         2,6,7, 1,8,6)
)
simpleNetwork(G, height="100px", width="100px", charge = 0)
```

## Week 3 

### Mon 2/15

* **Topics of the Day**: 

  * Euler's Formula for Planar Graphs
  * Euler paths
  * Bipartite Graphs

* Worksheets: 
  
  * Paths in Graphs (continued from last time): 
  [[PDF](graphs/paths-in-graphs.pdf)]
  [[HTML](graphs/paths-in-graphs.html)]
  
  * Bipartite Graphs (new): 
  [[PDF](graphs/bipartite-graphs.pdf)]
  [[HTML](graphs/bipartite-graphs.html)]
  
* Problem Set due each day this week

### Wed 2/17

* **Announcements**

  * Test 1 will be on Friday, March 5 (day after advising day)

* **Topics of the Day**

  * **Bipartite Graphs**: 
  [[PDF](graphs/bipartite-graphs.pdf)]
  [[HTML](graphs/bipartite-graphs.html)]
  
    * Complete this worksheet and be sure to check with me about the algorithm
  for determining if a graph is bipartite.
  
  * **Planar and Non-Planar Graphs**:
  [[PDF](graphs/planar-and-non-planar-graphs.pdf)]
  [[HTML]](graphs/planar-and-non-planar-graphs.html)
  
    * Work through question 3.
  
  * Check with me if your group has questions about any of the recent 
  graph things
  
### Fri 2/19

* **Test 1** will be two weeks from today.  More info next week.

  * [Google Doc](https://docs.google.com/document/d/1-vsGGT8mtMtXU9cPdLSW4YlpvSdw6c9XftEAL7Hy_B0/edit?usp=sharing) where you can post questions
  
    * I'll take a look at these before Monday and spend some time on them if needed
    before moving on to our next unit (Counting and Probability)
    
    * I will likely schedule a review session on March 4 (advising day).  That will
    be another opportunity to answer questions from this list.
    
    * I'll add additional sections next week.
    
* New Groups

```{r echo = FALSE, results = "asis", include = TRUE}
set.seed(2192021)
R <- shuffle(Roster) %>% group_by(section) %>% 
  mutate(group = rep(1:5, length.out = n())) %>%
  mutate(group = ifelse(section == "B", group + 5, group))
         
for(g in 1:10) {
  cat(paste0('  * Group ', g, ": ", paste(R %>% filter(group == g) %>% pull(name) %>% sort(), collapse = ", "), "\n"))
}
```

* **Planar and Non-Planar Graphs**:
  [[PDF](graphs/planar-and-non-planar-graphs.pdf)]
  [[HTML]](graphs/planar-and-non-planar-graphs.html)

## Week 4 

### Mon 2/22

* **Planar and Non-Planar Graphs**:
  [[PDF](graphs/planar-and-non-planar-graphs.pdf)]
  [[HTML]](graphs/planar-and-non-planar-graphs.html)
  
* Graphs Wrap-up

### Wed 2/24

* New Groups

```{r echo = FALSE, results = "asis", include = TRUE}
set.seed(2242021)
R <- shuffle(Roster) %>% group_by(section) %>% 
  mutate(group = rep(1:5, length.out = n())) %>%
  mutate(group = ifelse(section == "B", group + 5, group))
         
for(g in 1:10) {
  cat(paste0('  * Group ', g, ": ", paste(R %>% filter(group == g) %>% pull(name) %>% sort(), collapse = ", "), "\n"))
}
```


* **Topic of the day**: Counting without counting
  * Worksheet: 
  [[HTML]](counting-and-probability/some-counting-problems.html)
  [[PDF]](counting-and-probability/some-counting-problems.pdf)

### Fri 2/26

* [Test 1 Info](../test-info/index.html) is available on the test info page.

    * Review session next Thursday at noon (via Teams)

* [Homework](../hw/) notes

    * All homework due before Test 1 has been posted.
    * WW 4 is due on **Tuesday** next week (but you can do it sooner if you like).
    * I can't promise that the last problem set will be graded before the test (that 
    will depend on our graders' schedules), but you should have feedback from 
    everything else.
    
* Wrap-up from last time: 4 Counting Principles
    * Multiplication Principle
    * Addition Principle
    * Subtraction Principle
    * Inclusion-Exclusion

* **Topic of the day**: The Pigeonhole Principle

  * Worksheet: 
  [[HTML]](counting-and-probability/pigeonhole-principle.html)
  [[PDF]](counting-and-probability/pigeonhole-principle.pdf)

## Week 5 

### Mon 3/1

* Test on Friday
  * Review session Thursday at noon (via Teams)
 
* Homework due Monday, Tuesday, and Wednesday this week

* **Topic of the Day**: The Division Principle

  * Worksheet: 
  [[HTML]](counting-and-probability/division-principle.html)
  [[PDF]](counting-and-probability/division-principle.pdf)

### Wed 3/3

* Test on Friday
  * Review session Thursday at noon (via Teams)
    * Post questions [**here**](https://docs.google.com/document/d/1-vsGGT8mtMtXU9cPdLSW4YlpvSdw6c9XftEAL7Hy_B0/edit?usp=sharing)
  * Contact me TODAY if you are able to take the test in our classroom (NH 259) 
  at the regular time.  (Of course, if something like a COVID quarantine comes up, we'll
  adjust.)
  
* **Topic of the Day**: The Division (and other) Principles

  * More practice with counting problems
  * Binomial coefficients can help simplify your thought process
  * Worksheet: 
  [[HTML]](counting-and-probability/counting-problems.html)
  [[PDF]](counting-and-probability/counting-problems.pdf)

### Fri 3/5

* **Test 1**
  * In NH 259 (for most of you)

## Week 6 {.active}

### Mon 3/8

* Test have been graded; information will be posted later today.

* **Topic of the day**: Probability

  * Worksheet: [[HTML]](counting-and-probability/probability.html)
  