---
title: "Code Tidying"
author: "Math 252"
date: "5/7/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


Writing code that adheres to a style guide makes code maintenance easier
and is required by many companies and open source projects. Unfortunately,
people are not particularly good at following the rules.
Your task is to write a python [^1] program to do some code tidying.

[^1] If you want to use some other programming language, clear that with me 
first.

## Things to tidy

Here is a list of tasks for your tidyer.  Your job is to implement
as many as you can using regular expressions.

1. No trailing spaces at the ends of lines.

1. Tab characters at the beginning of a line should be replaced by (default 4) spaces.  Use `-t` to choose a different number.

1. Except at the beginning of a line (for indentation purposes), there should
be no double spaces.

1. Comment characters should be followed by a space. [Watch out for 
escaped symbols and string constants in quotes.]

    * Use `-c` or `--comment` to specify the comment characters.

1. All operators should be surrounded by spaces.  [Note: some operators like `==`, `>=`, etc. use more than one character.]

1. No space directly before a comma (`,`). Always a space after a comma.

1. There should be no spaces immediately following `(` or immediately 
preceeding `)`.

1. There should not be a space before an open paren [`(`] unless the previous word is `if`, `for`, or `while`.

1. If the flag `--python` is specified, for a more pythonic approach, 
convert `while (i < 3):` into `while i < 3:` since those parens are not required.  Do the same for `for` and `if`.


### Bonus features

1. Lines should be at most 79 characters long.  Break long lines in 
good places.

    * Use `-w` or `--width` to set the width to a different max.

1. Have an option `-r` or `--report` that only finds lines that do 
not comply with the style guide but does not correct them. 
Each offending line should be printed preceded by the line number.

    * double bonus: report the problem(s) as well as the line number


## Additional specifications

1. You program should comply with the tidy rules above.

2. Your program should executable at the command line and called `tidy`.

3. The following syntax should tidy the file `foo.py` and print the results 
on stdout.

```
tidy -f foo.py
```

4. If you add in extra features, use `-x` or `--extra` to turn them on.

```
tidy -f foo.py -x
```

## A stub

[Here](Python/tidy-template.py) is a stub you can start from.

```{r echo = FALSE, comment=''}
cat(readLines('Python/tidy-template.py'), sep = '\n')
```
