---
title: "Code Tidying"
author: "Math 252"
date: "5/7/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


Writing code that adheres to a style guide makes code maintenance easier
and is required by many companies and open source projects. Unfortunately,
people are not particularly good at following the rules.

Your task is to write a python program to do some code tidying.
(If you want to use some other programming language, clear that with me 
first.) This will be a purely syntactic tidying and is (mostly)
language agnostic (since most of the elements are a good idea in many
programming languages).  
You could, of course, fancy this up in many ways, including adding in
substitutions tailored for specific languages and auto-detecting
the langauge.

## Things to tidy

Here is a list of tasks for your `tidy` program. 
Your job is to implement as many as you can using regular expressions.

1. No trailing spaces at the ends of lines.

1. Tab characters at the beginning of a line should be replaced by (default 4) spaces.  Use `-t` to choose a different number.

1. Except at the beginning of a line (for indentation purposes), there should be no double spaces.  (In practice, this is probably too strict --
sometimes the use of space to align code makes it more readable.  Add
a flag to turn this off if you like, but leave it on by default.)

1. Comment characters should be followed by a space. [Watch out for 
escaped symbols and string constants in quotes.]

    * Use `-c` or `--comment` to specify the comment characters.

1. All operators should be surrounded by spaces.  [Note: some operators like `==`, `>=`, etc. use more than one character.]

1. No space directly before a comma. Always a space after a comma.

1. There should be no spaces immediately following `(` or immediately 
preceeding `)`.

1. There should not be a space before an open paren [`(`] unless the previous word is `if`, `for`, or `while`.

1. If the flag `--python` is specified, for a more pythonic approach, 
convert things like `while (i < 3):` into `while i < 3:` since those 
parens are not required.  Do the same for `for` and `if`.  You should 
be handle nested parentheses to some degree.


### Bonus features

1. Lines should be at most 79 characters long.  Break long lines in 
good places.

    * Use `-w` or `--width` to set the width to a different max.

1. Have an option `-r` or `--report` that only finds lines that do 
not comply with the style guide but does not correct them. 
Each offending line should be printed preceded by the line number.

    * double bonus: report the problem(s) as well as the line number

1. Add additional code tidying features if you like. You might consult
a style guide like PEP8 to get some ideas. Be sure to document them.

## Additional specifications

1. You program file should comply with the tidy rules above.

2. `tidy --help` should print useful help information.

3. Your program should be executable from the command line 
and called `tidy`.

4. The following syntax should tidy the file `foo.py` and print the results 
on stdout.

```
tidy -f foo.py
tidy -f foo.py --python   # with python bonus
```

4. If you add in extra features, use `-x` or `--extra` to turn them on.

```
tidy -f foo.py -x
```

5. You may create additional command line arguments as you see fit.

## Submission

Submit your program (named `tidy`) and any additional files 
to your `math252/regex` folder (just like you do for CS courses).

## A stub

[Here](Python/tidy-template.py) is a stub you can start from.

```{r echo = FALSE, comment=''}
cat(readLines('Python/tidy-template.py'), sep = '\n')
```
