---
title: "Probability Solutions"
author: "Math 252"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  comment = NA)
```


### 4: Urn

```{r}
choose(3, 2) +       # pick 2 red balls
  choose(5, 2) /     # or pick two blue balls
  choose(8, 2)       # pick 2 of the 8 balls
```

### 5: Dice

There are $6 \cdot 6 = 36$ possible equally likely rolls.  Four of them sum to 5, so 4/36.

### 6. Smaller Dice

$4/16$

### 7. 6 or Doubles

There are 6 ways to get doubles.  There are 5 ways to get a 6 ($1+5, 2+4, 3+3, 4+2, 5+1$), but notice that $3 + 3$
is in both lists, so we have $6 + 5 - 1$ ways to get a sum of six or doubles.

Answer; $10 / 36$

### 8. Lotto

Number of lotto tickets: $10^3 = 1000$.

Probabiliyt of big prize: $1/1000$.

Probability of small prize: $6/1000 - 1/1000 = 5/ 1000$. (There are $3 \cdot 2 \cdot 1$ ways to order three different numbers, but one of those
wins the big prize.)

If you pick 7-7-7, you can only win the big prize. (In the real lottery, your big prize is bigger to make up for 
not having a chance to win the smaller prize.)

### 9. Flush

```{r}
choose(4, 1) *       # pick a suit
  choose(13, 5) /    # pick 5 cards of that suit
  choose(52, 5)      # pick 5 cards from teh 52-card deck
```

### 10. 4-of-a-kind


```{r}
choose(13, 1) *      # pick a number to have 4 of
  choose(4, 4) *     # pick all four of that number
  choose(12, 1) *    # pick a number to have 1 of
  choose(4, 1) /     # pick 1 of those 4 cards
  choose(52, 5)      # pick 5 cards from teh 52-card deck
```

### 11, Random Bit String

Only one string doesn't have any 0's. So using the complement rule, we have 

```{r}
1 - 1 / 2^8
```

### 12. Yahtzee.

All numbers match:  

```{r}
6 / 6^5
```

At least 2 match = not all different, so

```{r}
1 - (6 * 5 * 4 * 3 * 2 * 1) / 6^5
```

### 13. Loaded Dice

If six is twice as likely as the rest, solve $ 5p + 2p = 1$ to get the probbility of rolling a 1.  $p - 1/7$.

Probability of rolling a 6 is $2/7$.

Probability of rolling doubles: $5 \cdot \frac17 \cdot \frac17 + \frac27 \cdot \frac27 = `r round(9/49, 4)`$

### 15. More Lottery

```{r}
# a. 
1 / choose(40, 6) 
choose(6,6) * choose(34, 0) / choose(40, 6) 

# b
choose(6, 5) * choose(34, 1) / choose(40, 6)
```

```{r echo = FALSE}
dplyr::tibble(correct = 6:0, 
              numerator = choose(6, correct) *choose(34, 6 - correct),
              probability = choose(6, correct) *choose(34, 6 - correct) / choose(40, 6)
) |>
  pander::pander()

```