---
title: "From Class"
author: ""
date: ""
output:
  html_document:
    includes:
      before_body: ../header.html
    css: ['../../styles/rpruim2020.css', '../../styles/folding.css']
---

```{r setup, include = FALSE}
library(tidyverse)
library(mosaic)
```

<script src="../../styles/folding.js"></script>


<style>
div.indent {
  margin: 0 0 0 25px;
}
</style>



## Worksheets/Handouts

I'll print and bring worksheets to class, but if you lose one or want to have it in
electronic format from some reason, you can get it here as well.

<!-- ### Models of Computation -->

<!-- * Worksheets:  -->
<!--   [PDF](models-of-computation/Models-of-Computation.pdf) -->
<!--   [HTML](models-of-computation) -->

<!-- ### Counting and Probability -->

<!-- * Worksheets:  -->
<!--   [PDF](counting-and-probability/Counting-and-Probability.pdf) -->
<!--   [HTML](counting-and-probability/) -->

<div class = "explain" label = "Before Test 1">
<div class = "indent">
<div class = "explain" label = "Groups starting 10 Jan">


```{r echo = FALSE, results = "asis", include = TRUE, message = FALSE}
set.seed(110)
n_groups <- function(x) {
  ifelse(x > 18, (x+3) %/% 4, x %/% 3)
}  

Roster <- read.csv('../../../webwork/Math252-S22-roster-clean.csv') %>%
  mutate(name = paste(first, last))

R <- 
  mosaic::shuffle(Roster) |>
  group_by(section) |>
  mutate(group = rep(1:(n_groups(n())), length.out = n())) |>
  mutate(group = paste0(section, group)) 
         
for(g in sort(unique(R$group))) {
  cat(paste0('  * Group ', g, ": ", paste(R %>% filter(group == g) %>% pull(name), collapse = ", "), "\n"))
}
```

</div>

<div class = "explain" label = "Groups starting 17 Jan">


```{r echo = FALSE, results = "asis", include = TRUE, message = FALSE}
set.seed(117)
n_groups <- function(x) {
  ifelse(x > 18, (x+3) %/% 4, x %/% 3)
}  

Roster <- read.csv('../../../webwork/Math252-S22-roster-clean2.csv') %>%
  mutate(name = paste(first, last))

R <- 
  mosaic::shuffle(Roster) |>
  group_by(section) |>
  mutate(group = rep(1:(n_groups(n())), length.out = n())) |>
  mutate(group = paste0(section, group)) 
         
for(g in sort(unique(R$group))) {
  cat(paste0('  * Group ', g, ": ", paste(R %>% filter(group == g) %>% pull(name), collapse = ", "), "\n"))
}
```

</div>

<div class = "explain" label = "Groups starting 24 Jan">

```{r echo = FALSE, results = "asis", include = TRUE, message = FALSE}
set.seed(124)
n_groups <- function(x) {
  ifelse(x > 18, (x+3) %/% 4, x %/% 3)
}  

Roster <- read.csv('../../../webwork/Math252-S22-roster-clean2.csv') %>%
  mutate(name = paste(first, last))

R <- 
  mosaic::shuffle(Roster) |>
  group_by(section) |>
  mutate(group = rep(1:(n_groups(n())), length.out = n())) |>
  mutate(group = paste0(section, group)) 
         
for(g in sort(unique(R$group))) {
  cat(paste0('  * Group ', g, ": ", paste(R %>% filter(group == g) %>% pull(name), collapse = ", "), "\n"))
}
```

</div>

<div class = "explain" label = "Groups starting 2 Feb">

```{r echo = FALSE, results = "asis", include = TRUE, message = FALSE}
set.seed(202)
n_groups <- function(x) {
  ifelse(x > 18, (x+3) %/% 4, x %/% 3)
}  

Roster <- read.csv('../../../webwork/Math252-S22-roster-clean2.csv') %>%
  mutate(name = paste(first, last))

R <- 
  mosaic::shuffle(Roster) |>
  group_by(section) |>
  mutate(group = rep(1:(n_groups(n())), length.out = n())) |>
  mutate(group = paste0(section, group)) 
         
for(g in sort(unique(R$group))) {
  cat(paste0('  * Group ', g, ": ", paste(R %>% filter(group == g) %>% pull(name), collapse = ", "), "\n"))
}
```

</div>

</div>


### Graphs

date     |  Topic                | link(s)
-------- | --------------------- | ------------------------------------------
Mon 1/10 | Intro to Graphs       | [PDF](graphs/intro-to-graphs.pdf)
Wed 1/12 | More Graphs           | [PDF](graphs/more-graphs.pdf)
Fri 1/14 | Graph Representations | [PDF](graphs/graph-representations.pdf)
Mon 1/17 | Graph Isomorphism     | [PDF](graphs/graph-isomorphism.pdf)
Wed 1/19 | Paths in Graphs       | [PDF](graphs/paths-in-graphs.pdf)
Fri 1/21 | Paths in Graphs       | [PDF](graphs/paths-in-graphs.pdf)
Mon 1/24 | Bipartite Graphs      | [PDF](graphs/bipartite-planar.pdf)
Wed 1/26 | Planar, Non-planar Graphs      | [PDF](graphs/planar-and-non-planar-graphs.pdf)

### Counting and Probability

date     |  Topic                | link(s)
-------- | --------------------- | ------------------------------------------
Mon 1/31 | Some Counting Problems | [PDF](counting-and-probability/some-counting-problems.pdf)
Wed 2/02 | Pigeon Hole Principle | [PDF](counting-and-probability/pigeonhole-principle.pdf)
Fri 2/04 | Division Principle    | [PDF](counting-and-probability/division-principle.pdf)
Mon 2/07 | Donuts                | [PDF](counting-and-probability/counting-problems.pdf)

</div>

<div class = "explain" label = "Between Test 1 and Test 2">

### Groups

We will often work in groups. Periodically we will shuffle things up so you get to
work with new people.

<div class = "explain" label = "Groups starting 11 Feb">


```{r echo = FALSE, results = "asis", include = TRUE, message = FALSE}
set.seed(1102)
n_groups <- function(x) {
  ifelse(x > 18, (x+3) %/% 4, x %/% 3)
}  

Roster <- read.csv('../../../webwork/Math252-S22-roster-clean2.csv') %>%
  mutate(name = paste(first, last))

R <- 
  mosaic::shuffle(Roster) |>
  group_by(section) |>
  mutate(group = rep(1:(n_groups(n())), length.out = n())) |>
  mutate(group = paste0(section, group)) 
         
for(g in sort(unique(R$group))) {
  cat(paste0('  * Group ', g, ": ", paste(R %>% filter(group == g) %>% pull(name), collapse = ", "), "\n"))
}
```

</div>
<div class = "explain" label = "Groups starting 21 Feb">

```{r echo = FALSE, results = "asis", include = TRUE, message = FALSE}
set.seed(221)
n_groups <- function(x) {
  ifelse(x > 18, (x+3) %/% 4, x %/% 3)
}  

Roster <- read.csv('../../../webwork/Math252-S22-roster-clean2.csv') %>%
  mutate(name = paste(first, last))

R <- 
  mosaic::shuffle(Roster) |>
  group_by(section) |>
  mutate(group = rep(1:(n_groups(n())), length.out = n())) |>
  mutate(group = paste0(section, group)) 
         
for(g in sort(unique(R$group))) {
  cat(paste0('  * Group ', g, ": ", paste(R %>% filter(group == g) %>% pull(name), collapse = ", "), "\n"))
}
```

</div>

<div class = "explain" label = "Groups starting 7 March">

```{r echo = FALSE, results = "asis", include = TRUE, message = FALSE}
set.seed(307)
n_groups <- function(x) {
  ifelse(x > 18, (x+3) %/% 4, x %/% 3)
}  

Roster <- read.csv('../../../webwork/Math252-S22-roster-clean2.csv') %>%
  mutate(name = paste(first, last))

R <- 
  mosaic::shuffle(Roster) |>
  group_by(section) |>
  mutate(group = rep(1:(n_groups(n())), length.out = n())) |>
  mutate(group = paste0(section, group)) 
         
for(g in sort(unique(R$group))) {
  cat(paste0('  * Group ', g, ": ", paste(R %>% filter(group == g) %>% pull(name), collapse = ", "), "\n"))
}
```

</div>

### Counting and Probability

date     |  Topic                | link(s)
-------- | --------------------- | ------------------------------------------
Mon 1/31 | Some Counting Problems | [PDF](counting-and-probability/some-counting-problems.pdf)
Wed 2/02 | Pigeon Hole Principle | [PDF](counting-and-probability/pigeonhole-principle.pdf)
Fri 2/04 | Division Principle    | [PDF](counting-and-probability/division-principle.pdf)
Mon 2/07 | Donuts                | [PDF](counting-and-probability/counting-problems.pdf)
Fri 2/11 | Probability           | [PDF](counting-and-probability/probability.pdf)
Mon 2/11 | Probability (cont'd)  | [PDF](counting-and-probability/probability.pdf)
Wed 2/16 | Conditional Probability 1 | [PDF](counting-and-probability/conditional-probability.pdf)
Fri 2/18 | Conditional Probability 2 | [PDF](counting-and-probability/probability-trees-and-bayes.pdf)
Mon 2/21 | Random Variables      | [PDF](counting-and-probability/random-variables.pdf)
Wed 2/23 | Expected Value        | [PDF](counting-and-probability/expected-value.pdf)
Fri 2/25 | More Expected Value   | [PDF](counting-and-probability/more-expected-value.pdf)
Mon 3/7  | Probability Wrap-Up   | [PDF](probability-wrapup.pdf)

  
### Langauges, Grammars, Machines

date     |  Topic                | link(s)
-------- | --------------------- | ------------------------------------------
Wed 3/9  | Grammars              | [PDF](models-of-computation/alphabets-languages-and-grammars.pdf)
Fri 3/11 | Types of Grammars     | [PDF](models-of-computation/types-of-grammars.pdf)

<!-- 05/07 | [Python regex project description](Python-tidy-project.html) -->
<!-- 05/07 | [Regular expressions in python](https://docs.google.com/presentation/d/1bjzMKrH8LeGmHC23fcX6z-gx3zW958VHTkjYyF1-Pdg/edit?usp=sharing) -->
<!-- 05/04 | [Turing Machine specification](http://www.calvin.edu/~rpruim/courses/m156/F16/TM/tmdoc.shtml) -->
<!-- 04/30 | [Turing Machine Simulator](http://www.6by9.net/z/projects/turingMachine/turingMachineApplet/) -->
<!-- 04/30 | [Turing Machines worksheet](Turing-Machines.pdf) -->
<!-- 04/27 | [DFA to regex conversion worksheet](DFA-to-regex.pdf) -->
<!-- 04/27 | [DFA to regex conversion notes](https://courses.engr.illinois.edu/cs374/fa2017/extra_notes/01_nfa_to_reg.pdf) -->
<!-- 04/25 | [Regular Languages Wrap-Up](Regular-WrapUp.pdf) -->
<!-- 04/23 | [Regular Grammars](Regular-Grammars.pdf) -->
<!-- 04/19 | [Regular Languages](RegularLanguages.pdf) -->
<!-- 04/14 | [Regular Languages](RegularLanguages.pdf) -->
<!-- 04/11 | [Automata](Automata.pdf) -->
<!-- 04/09 | [Automata](Automata.pdf) -->
<!-- 04/06 | [Grammars](Grammars.pdf) -->

</div>
<br>

------------


<!-- ## Other Resources -->

<!-- Date  | Resource -->
<!-- ----- | --------------------------------------------------- -->
<!-- 02/03 | [Wikipedia's page of graph terminology](https://en.wikipedia.org/wiki/Glossary_of_graph_theory_terms) -->
