---
output: 
  html_document:
    includes:
      before_body: ../../header.shtml
    css: ['../../../styles/rpruim2017a.css', '../../../styles/folding.css']
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

\def\prob{p}
\def\Prob{p}
\def\intersect{\cap}
\def\union{\cup}


<script src="../../../styles/folding.js"></script>

### Daily Dope Sheet 

## Wednesday, April 1 -- Expected Value

#### Homework 

* due Wednesday, April 1: WW 9
      
* due Friday, April 3: WW 10 and PS 8 

#### Stuff to do

1. Get today's worksheets. We'll be working mainly on Section 9 (Expected Value)

    * [PDF](../counting-and-probability/Counting-and-Probability.pdf) 
    * [HTML -- section 9](../counting-and-probability/expected-value.html)

2. Expected Value

    * The most important things today: 
        * Understanding how to calculate expected value for a discrete random variable
        * Understnading what expected value tells you about a random varaible.
    * Be sure to complete Exercise 1 - 12.  
    * Exercises 13 and 14 are bonus problems which are a little trickier.  But if things 
    are going well, you might like to give them a try.  (13 is easier than 14.)


#### Comments/Hints/Solutions

Check our MS Team channels as well.


<div class = "explain" label = "GPA">
<div class = "explain" label = "Exercise 9.1">
The student got more A's than C's, so the average should be higher than a B.
</div>

<div class = "explain" label = "Exercise 9.2">
Add up the 10 grades and divide by 10 -- just like you learned in grade school:
$$
\frac{4 + 4 + 4 + 4 + 4 + 3 + 3 + 3 + 3 + 2}{10}
$$
</div>

<div class = "explain" label = "Exercise 9.3">
\begin{align*}
\mbox{GPA} &= 
\frac{4 + 4 + 4 + 4 + 4 + 3 + 3 + 3 + 3 + 2}{10} \\
&=
\frac{5 \cdot 4}{10} + \frac{3 \cdot 4}{10} +\frac{1 \cdot 2}{10} \\
& =  4 \cdot \frac{\Large{5}}{10} + 3 \cdot \frac{\Large{4}}{10} + 2 \cdot \frac{\Large{1}}{10}
\end{align*}

</div>

<div class = "explain" label = "Exercise 9.4">

value         |   4   |   3   |   2   
-------------:|:-----:|:-----:|:-------
probability   |  5/10 |  4/10 |  1/10
</div>

<div class = "explain" label = "Exercise 9.5">
<video width="600" controls>
  <source src="recordings/gpa-expected-value.mp4" type="video/mp4">
  Your browser does not support HTML5 video.
</video>
</div>
</div>

<div class = "explain" label = "Exercise 9.6">
<video width="600" controls>
  <source src="recordings/exp-value-from-table.mp4" type="video/mp4">
  Your browser does not support HTML5 video.
</video>
</div>

<div class = "explain" label = "Exercise 9.7">
First we creat the probability table:

value         |   0   |   1   |   2    |   3
-------------:|:-----:|:-----:|:------:|:------
probability   |  1/8  |  3/8  |  3/8   |  1/8

From this we get the expected value as **"sum of values times probabilities"**:

$$
\mbox{expected value} = 0 \cdot \frac18 + 1 \cdot \frac38 + 2 \cdot \frac38 + 3 \cdot \frac18 = \frac{12}{8} = 1.5
$$
This makes sense: 1.5 is half of 3 and we should get half as many heads as flips (on average).
</div>

<div class = "explain" label = "Exercise 9.8">
The mathematical notation for "sum of values times probabilities" is 

$$
\sum_{x} x \cdot \Prob(X = x)
$$
where the sum is over all possible values of $x$.

If you don't like using both upper and lower case X (and keeping track of which goes where), you
could use a different letter, like this:

$$
\sum_{k} k \cdot \Prob(X = k)
$$
Now $k$ runs over all possible values of the random variable.

Or you can think in terms of pseudo code:

```
def expecte_value(values, prob): 
  ev = 0
  for v in values: 
    ev = ev + v * prob(v)
  return ev
```

If you work in R (a programming language for statistics), this is supper easy:

```{r}
values = c(0, 1, 2, 3)
probs = c(1/8, 3/8, 3/8, 1/8)
sum(values * probs)
```

</div>

<div class = "explain" label = "Exercise 9.9">
<video width="600" controls>
  <source src="recordings/exp-value-raffle.mp4" type="video/mp4">
  Your browser does not support HTML5 video.
</video>
</div>

<div class = "explain" label = "Exercise 9.10">
There are $4 \cdot 4 = 16$ equally likely rolls of the two dice.  We can get our 
probability table by counting up how many combinations give each difference.

value         |   0   |   1   |   2    |   3
-------------:|:-----:|:-----:|:------:|:------
probability   | 4/16  | 6/16  |  4/16  |  2/16

From here the expected value is (as always) the sum of values times probabilities.

expected value = $0 \cdot \frac{4}{16} + 1 \cdot \frac{6}{16} + 2 \cdot \frac{4}{16} + 3 \cdot \frac{2}{16}$

</div>

<div class = "explain" label = "Exercise 9.11">

Start by filling in this table

number of diamonds |   0   |   1   |   2    |   3   |   4
------------------:|:-----:|:-----:|:------:|:-----:|:------
probability        |   ??  |   ??  |   ??   |   ??  |  ??

Here's one example: $\displaystyle P(D = 2) = \frac{\binom{13}{2} \binom{39}{3}}{\binom{52}{5}}$

Fill in the rest of the tale and then "sum of values times probabilities".

```{r include = FALSE}
x = 0:4
ExpVal = sum( x * choose(13, x) * choose(39, 5-x) / choose(52, 5) )
```

You should get `r ExpVal`
</div>

<div class = "explain" label = "Exercise 9.12">

```{r include = FALSE}
vals <- 0:5
probs <- dbinom(vals, 5, 1/6)
exp_sixes <- sum(vals * probs)
f <- as.character(mosaic::fractions(exp_sixes))
```

Answer: `r exp_sixes` = `r f`.

Do you see why you should have expected that answer?
</div>

<div class = "explain" label = "Exercise 9.13">

```{r, include = FALSE}
vals <- 1:4
probs <- diff(c(0, choose(4,vals) * choose(vals*13, 5) / choose(52, 5))); probs
sum(probs)
exp_suits <- sum(vals * probs); exp_suits
```

Answer: `r exp_suits`.

Hint: Be carful not to overcount. Do your probabilities add up to 1?
</div>

<div class = "explain" label = "Exercise 9.14">
You will have to be especially careful to avoid (or adjust for) double counting here.
This is a challenging problem (given what we currently know).

Answer: `r 6 * (1 - (5/6)^5)`.

</div>

<br><br>