---
output: 
  html_document:
    includes:
      before_body: ../../header.shtml
    css: ../../../styles/rpruim2017a.css
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

\def\prob{p}
\def\intersect{\cap}
\def\union{\cup}

<style>
div.hide {
  display: none;
}
</style>

### Daily Dope Sheet 

## Monday, March 16

---


#### Homework 

  * due today: WW 6
  * due Wednesday: WW 7
  * due Friday: WW 8 and PS 7 (Note: you can start working on PS 7 already now.)

#### Stuff to do before 11:30 / 12:30

1. Make sure you are able to access our MS Team for the course. 

    * You should have gotten an invite to the team, I think.
    * [This link](https://teams.microsoft.com/l/team/19%3a076ad0b56a6146509163f69ae34a2740%40thread.tacv2/conversations?groupId=fd1cdf22-72c3-4d88-b309-1a0939c70ec0&tenantId=756349b9-0610-4b01-917b-2a4ac10df947) should also get you to our team.
    * I recommend that you install the appropriate MS Teams app for your
    device(s). That's a nicer way to work than in the browser. (There is a
    little icon in the far lower left corner of the browser that lets you
    install the app.  Or you can go to the appropriate online store for your
    device.)

2. Create at least one post in the Anything Goes channel.  

    * Since it is the Anything Goes channel, you may write anything you like. But it is 
    important that you post something. This is how I'll know you have access.
    
    * You may start a new post or reply to someone else. But if you reply, your post should
    be a reply and related to the previous poster's comment.
    
    * Feel free to show off your awesome MS Teams prowess if you like, but a simple one sentence post is all that is required.
    
    * If you have trouble getting into Teams, let me know. I'll help if I'm able.
    
3. Read the [Welcome to online Math 252](Welcome.html). 

    * Once you get to the frequently asked questions, you can scan the questions and only 
    read the answers if you have those questions. 
    (Several of those were answered in an email already as well.)
    
    * If you have additional questions, post them in the General channel in our MS Team.
   

#### At 11:30 / 12:30

1. If at all possible, be online at 11:30 (Section A) or 12:30 (Section B) with
MS Teams up and running. It's fine if you choose the time of the section you
don't usually attend if that works better for you.

2. Start working on the Stuff to Do. If you have questions, post in the appropriate channel of
our team.

3. Shortly after 11:30 and again shortly after 12:30, I may try to initiate a "meeting" (basically a video chat), mainly to test it out.

    * Be sure to mute yourself when you are not talking. That makes the audio much better.
    * If you are with other students in the same room, it is important that at most one of 
    you has a mic on.  Else we may get some terrible feedback.


#### Stuff to do 


<div>
1. Get today's worksheet (section 6). 

    * [PDF](../counting-and-probability/Counting-and-Probability.pdf) 
    * [HTML](../counting-and-probability/conditional-probability.html)

2. Introduction to Conditional Probability

    a. You may have noticed the following problem on the last worksheet:

        *  **Boys and Girls.** A family has two children.  One of them is a boy.  What is the probability that the other is a girl?
    
        Write down a what you think the answer should be.
    
    b. Then listen to this discussion of the problem (Sorry for the flickering; I'll
    try to get that figured out so we don't have flickering in every video.)
    
        * [part 1](recordings/2kids-p1.mp4)
        * [part 2](recordings/2kids-p2.mp4)
        * [part 3](recordings/2kids-p3.mp4)
        * [part 4](recordings/2kids-p4.mp4)
    
    c. With that introduction, work on **Exercises 1-9.** Pay careful attention to notation.
    (You can check your work below.)
    
3. Product Rule and Independence

    a. The product rule follows from the definition of conditional probability
    by a little simple algebra. Make sure you [see how](cond-prod-algebra.mp4).
    
    b. Continue working on the problems on the worksheet.
    
    c. After you get to the birthday problem, add your birthday to the list
    [here](https://docs.google.com/document/d/1LR6QY8uJhhblcm8aq57KXOixOjNrpD7OCe4deK9yADA/edit?usp=sharing) and we'll see if anyone in our class (both sections) shares a birthday.

#### Hints and Solutions


<button type = 'button' class = "collapsible">Exercises 6.1 - 6.9 (some)</button>
<div class = "content">

notation    | meaning | probability 1 | probability 2
----------- | ------- | ------------- | -----------------
$p(R)$      | probability that kid likes red best | $15/45 = 1/3$ |
$p(G)$      | probability that kid is a girl | $20/45$ |
$p(R \intersect G)$      | probability that kid is a girl who likes red | $8/45$ | 
$p(R \mid G)$      | among girls, what is the probability that a kid likes red | $8/20$ | $\frac{\prob(R\intersect G)}{\prob(G)} =  \frac{8/45}{20/45} = 8/20$
$p(\overline{G} \mid R)$      | among those who like red, what is the probability of being a boy | $7/15$ | $\frac{\prob(\overline{G}\intersect R)}{\prob(R)} =  \frac{8/45}{20/45} = 8/20 = \frac{7/45}{15/45} = 7/15$

</div>

<button type = 'button' class = "collapsible">Exercise 6.10</button>
<div class = "content">
We can rewrite $\prob(B \mid A) = P(B)$, as 
$\frac{\prob(A \intersect B)}{\prob(A)} = \prob(B)$

As little algera makes this $\frac{\prob(B \intersect A)}{\prob(B)} = \prob(A)$,
so 
$\prob(A \mid B) = P(A)$.

</div>

<button type = 'button' class = "collapsible">Exercise 6.11</button>
<div class = "content">
$\prob(B \mid A) = P(B)$, so $\prob(A \intersect B) = \prob(A) \cdot \prob(B)$
</div>
    
<button type = 'button' class = "collapsible">Exercise 6.12 (hint)</button>
<div class = "content">
Think of $\prob(A \intersect B \intersect C)$ as 
$\prob((A \intersect B) \intersect C)$ and apply the product rule twice.
</div>
    
<button type = 'button' class = "collapsible">Exercise 6.13</button>
<div class = "content">
$\prob(A \intersect B \intersect C \intersect D) = 
\prob(A) \cdot \prob(B \mid A) \cdot \prob(C \mid A \intersect B) \cdot 
\prob (D \mid A \intersect B \intersect C)$

Notice: Each term is the probability of one event given all the *previous* 
events.
</div>
    
<button type = 'button' class = "collapsible">Exercise 6.14</button>
<div class = "content">
a) $\frac{6 \cdot 1}{6 \cdot 6} = \frac{6}{36} = \frac{1}{6}$
b) Let's let F be "the first number is between 1 and 6" and let
   M be "the second number matches the first".  Then we can write
   this as
   
    $$
    \prob(F \intersect M)  = \prob( F ) \cdot p(M \mid F) = 1 \cdot \frac{1}{6} = \frac{1}{6}
    $$
</div>

<button type = 'button' class = "collapsible">Exercise 6.16</button>
<div class = "content">
Let's try a version with just three people. To get the probability that
at least two of them share a birthday, we will compute the probabilty 
that they all have different birthdays and subtract from 1.

To simplify notation, let let $N_i$ mean that person $i$ has a "new"
birthday (no one with a smaller index has that birthday).
So $N_1$ is always true.  $N_2$ will be true unless the first two
people have the same birthday.  When will $N_3$ be false?

```{r include = FALSE}
birthday <- function(n) {
  1 - prod((366: (366 - n + 1))/366)
}
```

\begin{align*}
 1 - \prob(N_1)
 \cdot
 \prob(N_2 \mid N_1)
 \cdot
 \prob(N_3 \mid N_2 \intersect N_1)\\
 &= 1 - 1 \cdot \frac{365}{366} \cdot \frac{364}{365} =
`r round(birthday(3), 4)`
\end{align*}

Note: The conditional probabilities are important here.  Why?

To see more details, look on page 461.
</div>

</div>


<script>
var coll = document.getElementsByClassName("collapsible");
var i;

for (i = 0; i < coll.length; i++) {
  coll[i].addEventListener("click", function() {
    this.classList.toggle("active");
    var content = this.nextElementSibling;
    if (content.style.display === "block") {
      content.style.display = "none";
    } else {
      content.style.display = "block";
    }
  });
}
</script>