---
pagetitle: 'M252: 5/04/2020'
output: 
  html_document:
    includes:
      before_body: ../../header.shtml
    css: ['../../../styles/rpruim2017a.css', '../../../styles/folding.css']
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggformula)
library(dplyr)
theme_set(theme_bw())
```

\def\prob{p}
\def\Prob{p}
\def\intersect{\cap}
\def\union{\cup}
\def\To{\Rightarrow}
\def\mybold#1{\boldsymbol{#1}}


<script src="../../../styles/folding.js"></script>

### Daily Dope Sheet 

## Monday, May 5: Languages that are not regular

#### Homework 

* Due Wednesday:  PS 17 (submit via [gradescope](https://gradescope.com))

* Wednesday's assignment is short.  If you want to get a head start for Wednesday's
class, read the first part of section 4.5 of our Zybook, up to and including
Participation Exercise 4.5.2: Identifying the next configuration of a Turing machine.
We will talk more about Turing Machines in class on Wednesday.

    
#### Worksheets

Models of Computation

* [HTML](../models-of-computation/a-language-that-is-not-regular)
* [PDF](../models-of-computation/Models-of-Computation.pdf)


<div class = "foldable" label = "COVID-19 Problem on Test 2"> 

About half of you had a moodle glitch on this problem -- sorry about that.

The question was suppose to give you each a probability that people are infected
and a group size and ask you the probability that at least one person in the group
is infected.

This can easily be calculated using the complement rule.  If $p$ is the probability
and $g$ is the group size, the probability we want is 


$$ 1 - (1-p)^g $$

For example, if $p$ is 10% and $g$ is 10, we get

```{r}
p = 0.10
g = 10
1 - (1-p)^g
```

Here's a plot showing the probability of at least one person infected in groups 
of various sizes and infection rates from 0% to 20%.


```{r, echo = FALSE}
library(ggformula)
library(dplyr)
expand.grid(
  p = seq(0, 0.20, by = 0.0025),
  g = 5:50
) %>% 
  mutate(
    prob = 1 - dbinom(0, g, p)
  ) %>%
  gf_raster(prob ~ p + g) %>%
  gf_labs(x = "probability of infection",
          y = "size of group",
          title = "Probability of at least one infected person in a group"
  ) %>% 
  gf_refine(scale_fill_viridis_b(n.breaks = 10, option = "C"))
```

</div>

<div class = "foldable" label = "A language that is not regular"> 

In addition to the in class discussion, you might like to look at page 885 of the text.
</div>

<div class = "foldable" label = "The Pumping Lemma"> 

<div class = "foldable" label = "Ideas behind the Pumping Lemma"> 

The basic idea used to show that $\{0^n1^n \ mid n = 0, 1, 2, ... \}$ is not regular
can be turned into a lemma that doesn't mention DFAs at all.  

Suppose $L$ is a regular language.  Then there must be a DFA $M$ that recognizes
$L$.  This machine has some number of states, let's call that number $N$.

Now suppose we have a long string $x$ with more symbols than $N$.  (We will write this 
$l(x) > N$.  We can break our string into three parts:

* before the loop: $u$
    * What can we say about the length of $u$?
    
* in the loop: $v$
    * What can we say about the length of $v$?
    
* after the loop: $w$.

Explain why each of these strings is in $L$ if and only if $x \in L$.

* $uvw$
* $uw$
* $uvvw$
* $uv^kw$

Putting this all together gives us the **Pumping Lemma**: 
</div>

<div class = "foldable" label = "The Pumping Lemma -- Statement"> 
If $L$ is a regular language, then there is a number $N$ such that for any
string $x$ with $l(x) > N$, we can write $x$ as

$$ 
x = uvw
$$

such that

* $l(u) < N$,
* $l(v) > 0$,
* $x \in L$ if and only if $uv^kw \in L$ for every natural number $k$.

Important things to note:

* The Pumping Lemma says something about *long* strings.  There is some length
beyond which the $x = uvw$ splitting is guaranteed. It may or may not work for shorter
strings.

* You do not get to choose $u$, $v$, and $w$.  The Pumping Lemma tells you that
there is a way to split up $x$, not that every way of splitting up $x$ will work.

* The length constraints on $u$ and $v$ are often important in applications.
In particular, $l(uv) \le N$.


</div>

<div class = "foldable" label = "Pumping Lemma -- Example Application"> 

Let's try using the Pumping Lemma to show that ${0^n1^n \mid n = 0, 1, \dots}$ is not
regular.

Suppose $L$ were regular.
Let $N$ be the number given by the Pumping Lemma.  Consider $x = 0^N1^N$. 

* There is some way to split up $x$ as $x = uvw$ (because $x$ is long enough), and 
* $u$ and $v$ consist entirely of 0's, since $l(uw) \le N$.

Now consider the string $uw$.  $uw$ must be in $L$, but $uw$ has fewer 0's than 1's, so
it can't be in $L$.  This means our assumption that $L$ was regular must be false.

**Note:** Anything you can prove using the Pumping Lemma, you can also prove by going
back to the underlying argument about repeated states (pigeon hole principle), loops,
etc.  So you can think of the Pumping Lemma as an (optional) alternative way to think
about these problems.  I don't care which approach you take, as long as you do it correctly.
Any problem that says "by the Pumping Lemma" may also be done from the underlying argument.
</div>
</div>

