---
pagetitle: 'M252: 4/27/2020'
output: 
  html_document:
    includes:
      before_body: ../../header.shtml
    css: ['../../../styles/rpruim2017a.css', '../../../styles/folding.css']
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

\def\prob{p}
\def\Prob{p}
\def\intersect{\cap}
\def\union{\cup}
\def\To{\Rightarrow}
\def\mybold#1{\boldsymbol{#1}}


<script src="../../../styles/folding.js"></script>

### Daily Dope Sheet 

## Monday, April 27: DFA -> Regular Expression 

#### Homework 

* due Today: PS 16 (submit via [gradescope](https://gradescope.com))
* due Wednesday: no homework -- review for exam
    * Note: Rosen does not cover today's topic, so there are no exercises
    about converting from DFA or NFA to regular expression, but you are 
    expected to know how to do this for the test.
    
#### Wednesday is Review Day

I'll be in our Worksheets channel from 11:30 until people aren't around (but not past 1:30) to
answer questions in preparation for the test on Friday.

#### What we are learning today

Today we add the last arrow to this chart: DFA $\to$ regular expression.

$$
\Large
\begin{array}{ccc}
\mathrm{DFA} & \stackrel{\longrightarrow}\longleftarrow & \mathrm{NFA} \\
\downarrow& \nearrow & \downarrow \uparrow \\
\mathrm{regular}& & \mathrm{regular} \\
\mathrm{expression}& & \mathrm{grammar} 
\end{array}
$$

This shows that all four of our representations (DFA, NFA, regular grammar, regular expression)
are equivalent.  The languages recognized/generated/expressed
by DFAs/NFAs/regular grammars/regular expressions are called **regular langauges**.


#### Worksheet

We will cover Section 6.

* [HTML](../models-of-computation/regular-langauges.html)
* [PDF](../models-of-computation/Models-of-Computation.pdf)


#### Comments/Hints/Solutions

\def\emptystring{\lambda}
\def\union{\cup}



<div class = "foldable" label = "Exercise 6.2.1">

```{r echo = FALSE, fig.align = 'center'}
# path <- glue::glue('../images/DFA2regex1/frame_0000{i}.png', i = 1:9)
path <- dir('../images/DFA2regex1', full.names = TRUE)
knitr::include_graphics(path)
```

</div>

<div class = "foldable" label = "Exercise 6.2.2">

```{r echo = FALSE, fig.align = 'center'}
path <- dir('../images/DFA2regex2', full.names = TRUE)
knitr::include_graphics(path)
```

</div>
<div class = "foldable" label = "Exercise 6.2.3">

```{r echo = FALSE, fig.align = 'center'}
path <- dir('../images/DFA2regex3', full.names = TRUE)
knitr::include_graphics(path)
```

</div>

<div class = "foldable" label = "Exercise 6.2.4-5">
The algorithm works just the same if you start from an NFA or a GNFA.

As a side note, this implies that GNFAs are equivalent to NFAs.  In particular, $\lambda$-NFAs (NFAs that allow transitions
labeled $\lambda$ that can be used without processing any input symbols) are equivalent.  So our diagram looks like
this 

$$
\Large
\begin{array}{ccccccc}
\mathrm{DFA} & \longleftrightarrow & \mathrm{NFA} & \longrightarrow & \lambda\mathrm{-NFA} & \longrightarrow & \mathrm{GNFA} \\
&& \downarrow \uparrow & \nwarrow  &  & \swarrow  \\
& & \mathrm{regular}& & \mathrm{regular} \\
& & \mathrm{grammar}& & \mathrm{expression} 
\end{array}
$$
Some books and online resources will demonstrate the conversion 
$$
\mbox{regular expression} \longrightarrow \lambda\mathrm{-NFA}
$$
since this is slightly easier than 
$$
\mbox{regular expression} \longrightarrow \mathrm{NFA}
$$
But then we would need an additional conversion
$$
\lambda\mathrm{-NFA} \longrightarrow  \mathrm{NFA} \;.
$$
This doable, but isn't really any easier than building it into the 
the conversion from regular expressions to NFAs, the way we did it.

You should know how to perform all of the conversions illustrated in the diagram
above.
</div>

<div class = "foldable" label = "Exercise 6.3.1">
Compare your descriptions to the descriptions below.  Did you leave out 
anything important?  Do the solutions below leave out anything important?

1a. regex $\to$ NFA: 

* 6 parts: 
    * 3 base cases -- pretty easy
    * 3 that show union, concatenation and Kleene star -- these are the "interesting" part
  
* union: $A \cup B$ where $A = L(N_1)$ and $B = L(N_2)$.

    * keep all states and transitions from $N_1$ and $N_2$
    
    * add one new start state $S$; $S$ is final if either 
      start state of $N_1$ or start state of $N_2$ is final
      
    * add some additional transitions
        * if $S_1 \stackrel{x}{\to} A$, add $S \stackrel{x}{\to} A$
        * if $S_2 \stackrel{x}{\to} B$, add $S \stackrel{x}{\to} B$

* concatenation: $AB$ where $A = L(N_1)$ and $B = L(N_2)$.

    * keep all states and transitions from $N_1$ and $N_2$
    * start state from $N_1$ is the start state 
    * only final states from $N_2$ are final
    * add some additional transitions
        * if $A \stackrel{x}{\to} F$ (a final state in $N_1$), then add
          add $A \stackrel{x}{\to} S_2$ (start state in $N_2$)
          
* Kleene star: $A = L(N_1)$

    * keep all states and transitions from $N_1$ 
    * add a new start state that is a final state (because $\lambda \in A^*$)
    * add new transitions from new start state ($S$)
    
        * if $S_1 \stackrel{x}{\to} A$, add $S \stackrel{x}{\to} A$ (includes case where $A = S_1$)
        
    * add new transitions to old start state ($S_1$)
    
        * if $A \stackrel{x}{\to} F$, where F is final, add $A \stackrel{x}{\to} S_1$
        * If we are at the end of a string in $L(N_1)$, this lets us "restart"
    
        
1b. DFA $\to$ NFA

* easy: a DFA "is" an NFA

1c. NFA $\to$ DFA

* DFA states are **subsets of NFA states** (think: magnets on white board)

* A is final in DFA if A contains a final state of NFA 

* Remember that $\{ \}$ may be needed as one of the states in the DFA
    
1d. DFA (or NFA) $\to$ reg grammar

* $A \stackrel{x}{\to} B$ becomes 

    * $A \to xB$
    
    * also $A \to x$  if B is final state
    
* If S is final in NFA/DFA, then add $S \to \lambda$
    
* Number of rules = number of transitions + number of transitions to final states +
    1 more if the start state is final
    
1e. reg grammar $\to$ NFA

* states: non-terminals of grammar + 
one additional state $F$ (a final state)
    * Start symbol is also start state
    * If $S \to \lambda$ is a rule, then $S$ is a final state

* production rules become transitions in NFA

    * $A \to xB$ becomes $A \stackrel{x}{\to} B$
    * $A \to x$  becomes $A \stackrel{x}{\to} F$ (the added final state)
  
1f. DFA $\to$ regular expression
 
* Big idea: Make GNFAs with fewer and fewer states until we have a GNFA with two states 
(one the start state and the other the accepting state) and one transition between them, labeled
with a regular expression.

* If necessary, make two addjustments to the DFA before beginning
    * Make sure there is a lonely start state 
    * Make sure there is a single accepting state with no exiting transitions.
    * $\lambda$-transitions are allowed for this step, so we might not have a DFA anymore after these adjustments.
    
* Pick any internal state $R$ (not the start or accepting state) and "rip it out".
    * create or update transitions $A \to B$ when there was $A \to R \to B$.

</div>

<br><br>