---
pagetitle: M252 | from class
toc-depth: 3
---

# From Class

```{r setup, include = FALSE}
library(tidyverse)
library(mosaic)
```

<script src="styles/folding.js"></script>

<style>
div.indent {
  margin: 0 0 0 25px;
}
</style>


## Worksheets/Handouts

I'll print and bring worksheets to class, but if you lose one or want to have it in
electronic format for some reason, you can get it here as well.

<!-- ### Models of Computation -->

<!-- * Worksheets:  -->
<!--   [PDF](models-of-computation/Models-of-Computation.pdf) -->
<!--   [HTML](models-of-computation) -->

<!-- ### Counting and Probability -->

<!-- * Worksheets:  -->
<!--   [PDF](counting-and-probability/Counting-and-Probability.pdf) -->
<!--   [HTML](counting-and-probability/) -->

:::{.callout-note collapse="true"}
### Before Test 1

::::{.callout-note collapse="true"}
#### Groups starting 9 Jan

```{r echo = FALSE, results = "asis", include = TRUE, message = FALSE}
set.seed(109)
n_groups <- function(x) {
  ifelse(x > 18, (x+3) %/% 4, x %/% 3)
}  

Roster <- readxl::read_excel("../grades/S23/Rosters-Spring2023.xlsx") |>
  rename(first = `Preferred First Name`) |>
  rename(last = `Last Name`) |>
  rename(email = `Primary Institutional Email Address`) |>
  rename(name = `Student`) |>
  rename(id = `Student ID`) |>
  separate(`Course Section`, into = c('course', 'section', 'title'), sep = "\\s*-\\s*", remove = FALSE) |>
  mutate(short_name = paste(first, substr(last, 1, 1))) |>
  filter(course == "MATH 252")


R <- 
  mosaic::shuffle(Roster) |>
  group_by(section) |>
  mutate(group = rep(1:(n_groups(n())), length.out = n())) |>
  mutate(group = paste0(section, group)) 
         
for(g in sort(unique(R$group))) {
  cat(paste0('  * Group ', g, ": ", paste(R %>% filter(group == g) %>% pull(short_name), collapse = ", "), "\n"))
}
```

::::

::::{.callout-note collapse="true"}
#### Groups starting 16 Jan

```{r echo = FALSE, results = "asis", include = TRUE, message = FALSE}
set.seed(116)
n_groups <- function(x) {
  ifelse(x > 18, (x+3) %/% 4, x %/% 3)
}  

Roster <- readxl::read_excel("../grades/S23/Rosters-Spring2023.xlsx") |>
  rename(first = `Preferred First Name`) |>
  rename(last = `Last Name`) |>
  rename(email = `Primary Institutional Email Address`) |>
  rename(name = `Student`) |>
  rename(id = `Student ID`) |>
  separate(`Course Section`, into = c('course', 'section', 'title'), sep = "\\s*-\\s*", remove = FALSE) |>
  mutate(short_name = paste(first, substr(last, 1, 1))) |>
  filter(course == "MATH 252")


R <- 
  mosaic::shuffle(Roster) |>
  group_by(section) |>
  mutate(group = rep(1:(n_groups(n())), length.out = n())) |>
  mutate(group = paste0(section, group)) 
         
for(g in sort(unique(R$group))) {
  cat(paste0('  * Group ', g, ": ", paste(R %>% filter(group == g) %>% pull(short_name), collapse = ", "), "\n"))
}
```

::::

::::{.callout-note collapse="true"}
#### Groups starting 25 Jan

```{r echo = FALSE, results = "asis", include = TRUE, message = FALSE}
set.seed(125)
n_groups <- function(x) {
  ifelse(x > 18, (x+3) %/% 4, x %/% 3)
}  

R <- 
  mosaic::shuffle(Roster) |>
  group_by(section) |>
  mutate(group = rep(1:(n_groups(n())), length.out = n())) |>
  mutate(group = paste0(section, group)) 
         
for(g in sort(unique(R$group))) {
  cat(paste0('  * Group ', g, ": ", paste(R %>% filter(group == g) %>% pull(short_name), collapse = ", "), "\n"))
}
```

::::


::::{.callout-note collapse="true"}
#### Groups starting 01 Feb

```{r echo = FALSE, results = "asis", include = TRUE, message = FALSE}
set.seed(201)
n_groups <- function(x) {
  ifelse(x > 18, (x+3) %/% 4, x %/% 3)
}  

R <- 
  mosaic::shuffle(Roster) |>
  group_by(section) |>
  mutate(group = rep(1:(n_groups(n())), length.out = n())) |>
  mutate(group = paste0(section, group)) 
         
for(g in sort(unique(R$group))) {
  cat(paste0('  * Group ', g, ": ", paste(R %>% filter(group == g) %>% pull(short_name), collapse = ", "), "\n"))
}
```
::::



### Graphs

```{block}
date     |  Topic                | link(s)
-------- | --------------------- | ------------------------------------------
Mon 1/9  | Intro to Graphs       | [PDF](/from-class/01-graphs.pdf)
Wed 1/11 | More Graphs           | [PDF](/from-class/02-graphs.pdf)
Wed 1/13 | Representing Graphs   | [PDF](/from-class/03-graphs.pdf)
Mon 1/16 | Graph Isomorphism     | [PDF](/from-class/04-graphs.pdf)
Wed 1/18 | Paths in Graphs       | [PDF](/from-class/05-graphs.pdf)
Fri 1/20 | Paths in Graphs       | [PDF](/from-class/05-graphs.pdf)
Mon 1/23 | Bipartite Graphs      | [PDF](/from-class/06-graphs.pdf)
Wed 1/25 | Planar Graphs         | [PDF](/from-class/07-graphs.pdf)
```

```{block}
#| include: true
### Counting and Probability

date     |  Topic                | link(s)
-------- | --------------------- | ------------------------------------------
Mon 1/30 | Some Counting Problems | [PDF](/from-class/08-counting.pdf)
Mon 1/30 | Some Counting Problems (solutions) | [PDF](/from-class/08-counting-sols.pdf)
Wed 2/01 | Pigeon Hole Principle | [PDF](/from-class/09-counting.pdf)
Fri 2/03 | Division Principle    | [PDF](/from-class/10-counting.pdf)
Mon 2/06 | Donuts                | [PDF](/from-class/11-counting.pdf)
```

:::


:::{.callout-note collapse="true"}
### Between Test 1 and Test 2

::::{.callout-note collapse="true"}
#### Groups starting 10 Feb

```{r echo = FALSE, results = "asis", include = TRUE, message = FALSE}
set.seed(210)
n_groups <- function(x) {
  ifelse(x > 18, (x+3) %/% 4, x %/% 3)
}  

R <- 
  mosaic::shuffle(Roster) |>
  group_by(section) |>
  mutate(group = rep(1:(n_groups(n())), length.out = n())) |>
  mutate(group = paste0(section, group)) 
         
for(g in sort(unique(R$group))) {
  cat(paste0('  * Group ', g, ": ", paste(R %>% filter(group == g) %>% pull(short_name), collapse = ", "), "\n"))
}
```
::::

::::{.callout-note collapse="true"}
#### Groups starting 20 Feb

```{r echo = FALSE, results = "asis", include = TRUE, message = FALSE}
set.seed(220)
n_groups <- function(x) {
  ifelse(x > 18, (x+3) %/% 4, x %/% 3)
}  

R <- 
  mosaic::shuffle(Roster) |>
  group_by(section) |>
  mutate(group = rep(1:(n_groups(n())), length.out = n())) |>
  mutate(group = paste0(section, group)) 
         
for(g in sort(unique(R$group))) {
  cat(paste0('  * Group ', g, ": ", paste(R %>% filter(group == g) %>% pull(short_name), collapse = ", "), "\n"))
}
```
::::

::::{.callout-note collapse="true"}
#### Groups starting 6 March

```{r echo = FALSE, results = "asis", include = TRUE, message = FALSE}
set.seed(306)
n_groups <- function(x) {
  ifelse(x > 18, (x+3) %/% 4, x %/% 3)
}  

R <- 
  mosaic::shuffle(Roster) |>
  group_by(section) |>
  mutate(group = rep(1:(n_groups(n())), length.out = n())) |>
  mutate(group = paste0(section, group)) 
         
for(g in sort(unique(R$group))) {
  cat(paste0('  * Group ', g, ": ", paste(R %>% filter(group == g) %>% pull(short_name), collapse = ", "), "\n"))
}
```
::::

```{block}
#| include: true
### Counting and Probability

date     |  Topic                  | link(s)
-------- | ----------------------- | ------------------------------------------
Fri 2/10 | Intro to Probability    | [PDF](/from-class/12-probability.pdf)
Mon 2/13 | Intro to Probability    | [PDF](/from-class/12-probability.pdf) [solutions](/from-class/12-probability-sols.pdf)
Wed 2/15 | Conditional Probability | [PDF](/from-class/13-probability.pdf)
Fri 2/17 | More Conditional Probability | [PDF](/from-class/14-probability.pdf)
Mon 2/20 | Random Variables        | [PDF](/from-class/15-probability.pdf)
Mon 2/24 | Expected Value          | [PDF](/from-class/16-probability.pdf)
Mon 3/6  | Expected Value          | [PDF](/from-class/17-probability.pdf)
Wed 3/8  | Probability Review      | [PDF](/from-class/18-probability.pdf)
Fri 3/10 | Probability Review Solutions  | [PDF](/from-class/18-probability-sols.pdf)
Fri 3/10 | Grammars                | [PDF](/from-class/19-grammars.pdf)
```

:::


:::{.callout-note collapse="false"}

### After Test 2 

::::{.callout-note collapse="true"}
#### Groups starting 15 March

```{r echo = FALSE, results = "asis", include = TRUE, message = FALSE}
set.seed(315)
n_groups <- function(x) {
  ifelse(x > 18, (x+3) %/% 4, x %/% 3)
}  

R <- 
  mosaic::shuffle(Roster) |>
  group_by(section) |>
  mutate(group = rep(1:(n_groups(n())), length.out = n())) |>
  mutate(group = paste0(section, group)) 
         
for(g in sort(unique(R$group))) {
  cat(paste0('  * Group ', g, ": ", paste(R %>% filter(group == g) %>% pull(short_name), collapse = ", "), "\n"))
}
```
::::

::::{.callout-note collapse="true"}

#### Groups starting 27 March

```{r echo = FALSE, results = "asis", include = TRUE, message = FALSE}
set.seed(324)
n_groups <- function(x) {
  ifelse(x > 18, (x+3) %/% 4, x %/% 3)
}  

R <- 
  mosaic::shuffle(Roster) |>
  group_by(section) |>
  mutate(group = rep(1:(n_groups(n())), length.out = n())) |>
  mutate(group = paste0(section, group)) 
         
for(g in sort(unique(R$group))) {
  cat(paste0('  * Group ', g, ": ", paste(R %>% filter(group == g) %>% pull(short_name), collapse = ", "), "\n"))
}
```

::::

::::{.callout-note collapse="true"}

#### Groups starting 3 April

Pick your own group this week!
::::


```{block}
#| include: true
### Counting and Probability

date     |  Topic                  | link(s)
-------- | ----------------------- | ------------------------------------------
Fri 3/10 | Grammars                | [PDF](/from-class/19-grammars.pdf)
Wed 3/15 | Types of Grammars       | [PDF](/from-class/20-grammars.pdf)
Fri 3/17 | Automata                | [PDF](/from-class/21-dfa.pdf)
Mon 3/20 | Non-determinism         | [PDF](/from-class/21a-nfa.pdf)
Fri 3/24 | Non-determinism         | [PDF](/from-class/21a-nfa.pdf)
Mon 3/27 | Regular Expressions     | [PDF](/from-class/22a-regex.pdf)
Wed 3/29 | Regular Expressions     | [PDF](/from-class/22b-regex.pdf)
Fri 3/31 | Regular Grammars        | [PDF](/from-class/23-grammars-and-nfas.pdf)
Mon 4/3  | NFA -> Regex            | [PDF](/from-class/24-nfa-to-regex.pdf)
Fri 4/14 | Turing Machines         | [PDF](/from-class/26-TM.pdf)
Mon 4/17 | More Turing Machines.   | [PDF](/from-class/27-TM.pdf)
Mon 4/17 | A Language that isn't regular | [PDF](/from-class/25-pumping.pdf)
Wed 4/19 | Python Regular Expressions | [slides](https://docs.google.com/presentation/d/1bjzMKrH8LeGmHC23fcX6z-gx3zW958VHTkjYyF1-Pdg/edit?usp=sharing)
```

::::{.callout-note collapse=true}

#### Regex games, puzzles, etc.

Note: You don't have to turn in any homework for Python regular expressions, but you should 
work through some of these games and tutorials to hone your skills. 
*This topic will be  included on the final exam.*

* [The regular expression game](http://play.inginf.units.it/#/)
    * Construct regular expressions to match the indicated portions of a page of text.
    * These tasks seem to mimic the kinds of things you might really do (finding authors
    in a bibilography, finding MAC addresses in some text, etc.)
    

* [regex 101](https://regex101.com/) 

    * [regex 101 quiz](https://regex101.com/quiz)

* [Regex crossword game](https://regexcrossword.com/)
    * Fill in the crossword with text that matches regex clues.
    * This forces you to work backwards, the regular expression are given
    
* [Regular expression golf](https://alf.nu/RegexGolf)
    * Find a short regex that matches everything in one column but doesn't match
    any portion of anything in the other column.
    * Some of the hard ones here get pretty hard and not very "natural"
    * Least useful of the three for learning regex.
::::

* [Some comments, hints, and solutions](from-class/models-of-computation-sols.qmd)

:::
