---
output:
  html_document:
    includes:
      before_body: ../header.shtml
    css: ../../styles/rpruim2017.css
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
require(dplyr)
require(tidyr)
require(kable)
require(kableExtra)
require(scales)
t1bks <- c(20, 33, 45, 57, 69, 83) %>% rev()
  # c(20, 43, 54, 65, 76, 88) %>% rev()
t2bks <- c(30, 55, 65, 75, 85, 96) %>% rev()
t3bks <- c(15, 36, 44, 52, 61, 72)   %>% rev()
fbks  <- c(25, 74, 96, 118, 140, 165)  %>% rev()
GS <- 
  data_frame(
    Grade = LETTERS[1:4],
    `test 1` = head(paste(lead(t1bks), "-", (t1bks)), 4)
    # `test 2` = head(paste(lead(t2bks), "-", (t2bks)), 4),
    # `test 3` = head(paste(lead(t3bks), "-", (t3bks)), 4),
    # `final`  = head(paste(lead(fbks),  "-", (fbks)), 4)
  ) %>% 
  arrange(Grade)
```

## Test Information

Information about tests will be posted here shortly before each test.

### Black out date requests

You can request that we not have a test on a particular day by filling out 
[this form](https://goo.gl/forms/tspWxJSW0K9Qa4gv1). I can't promise to 
meet all requests, but I'll do what I can to work with the data.

### Grading Scales

Here are the grading scales used for each of our tests and the final.

```{r, echo = FALSE}
kable(GS) %>%
  kable_styling(bootstrap_options = "striped", full_width = F) %>%
  row_spec(1, bold = F, color = "gray20", background = alpha("steelblue", 0.3)) %>%
  row_spec(2, bold = F, color = "gray20", background = alpha("navy", 0.3)) %>%
  row_spec(4, bold = F, color = "gray20", background = alpha("forestgreen", 0.3)) %>%
  row_spec(3, bold = F, color = "gray20", background = alpha("limegreen", 0.2))
```

### Test 1

**Date:** Wed, March 6

**Material Covered**: Graphs (Rosen chapter 10.1-10.5, 10.7) and Counting (Rosen 6.1-6.5)

#### Topics

**Graphs:**
modeling with graphs (what do edges and vertices represent?);
types of graphs(directed, undirected, self-loops, etc., etc.);
graph terminology (examples: vertex, edge, adjacent, degree, bipartite, path, connected,
connected component);
graph representations (adjacency matrix, incidence matrix, edge list, adjacency list);
graph isomorphism; graph invariants;
Euler paths and circuits; Hamilton paths and circuits;
planar graphs; Euler's formula; Kuratowski's Theorem

**Counting:**
bijection principle (no double counting, no skipping); 
multiplication principle; addition principle;  subtraction principle;
division principle; inclusion-exclusion; 
pigeonhole principle;
permutations and combinations; 
binomial coefficients.


### Test 2

**Date:** Wed, April 3

**Material Covered**: Counting and Probability (Rosen 6 and 7)

#### Topics


**Counting:**
Bijection principle (no double counting, no skipping);
multiplication principle;
addition principle;  division principle;
inclusion-exclusion; 
permutations and combinations; binomial coefficients;
inclusion-exclusion.

**Probability:**
Definition of finite probability and applications involving counting problems;
empirical vs theoretical probability;
3 probability axioms;
complement rule;
equally likely rule;
union rule;
conditional probability;
indepedence;
Bayes probability (e.g., false positives and false negatives);
probability trees;
random variables;
expected value of a random variable.

Note: Counting topics will largely arise in the context of probability calculations.

<!-- ### Test 3  -->

<!-- **Date:** Wed, May 2 -->

<!-- **Material Covered**: Languages, Grammars, and Automata -->
<!-- (Rosen 12.1 -- 12.4) -->

<!-- #### Topics -->

<!-- **Grammars:** components of a grammar (terminals, nonterminals, start symbol, -->
<!-- production rules, etc.), derivations, language generated by a grammar, types of -->
<!-- grammars (regular, context sensitive, context free). -->

<!-- **Automata:** components of an automaton (input alphabet, states, start state, -->
<!-- final states, transition funciton), DFA, NFA, GNFA, language recognized by  -->
<!-- an automaton, building up a complex automaton from component -->
<!-- automata. -->

<!-- **Regular expressions:** notation, 3 base cases and three operations, language -->
<!-- represented by a regular expression. -->

<!-- **Regular languages:** Equivalence of regular expressions, regular grammars, -->
<!-- DFAs, and NFAs.  (You should know how to do all of the conversions we have -->
<!-- learned.)  Pumping Lemma and showing languages are not regular. -->

